---
title: "ADS503_Team_Project"
author: "Team 6"
date: "6/7/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(mlbench)
library(Hmisc)
library(e1071)
library(caret)
library(tidyr)
library(corrplot)
library(AppliedPredictiveModeling)
library(car)
library(lattice)
library(lars)
library(stats)
library(pls)
library(dplyr)
library(kernlab)
library(randomForest)
library(gbm)
library(earth)
library(plotmo)
library(plotrix)
library(TeachingDemos)

```

#### Data Setup/Load

```{r setup_part2}

brain=read.table(file.choose(), header=TRUE, sep=",")
#RF_test_data=read.table(file.choose(), header=TRUE, sep=",")

```

```{r exploredata}

dim(brain_df)

head(brain)
brain[,2]

```
#### Histograms, EDA, observe Skew/etc 


```{r moredataexplore1}
### Remove low variance predictors:
### Does not work well with this data...
#brain_idx <- nearZeroVar(brain_df[,-c(1,2)]) 
#excluding the dependent variable
#brain_df_var <- brain_df[,-brain_idx]
#dim(brain_df_var)

hist(brain$X1007_s_at)
hist(brain$X1053_at)
hist(brain$X117_at)
hist(brain$X121_at)
```

***
*** 

### Logistic Regression

#### Generate binary version of output and add to dataframe **brain_df**

```{r dataprocesses1}
### Create a binary alternate output and include in dataframe

#Create an alternate output variable called Cancer with "Yes" or "No" 
#binary output so as to facilitate an odds ratio output from a Logistic Classifier
cancer2=brain$type
cancer2 = as.character(brain$type)
cancer2[cancer2 == "ependymoma"] = "Yes"
cancer2[cancer2 == "glioblastoma"] = "Yes"
cancer2[cancer2 == "medulloblastoma"] = "Yes"
cancer2[cancer2 == "pilocytic_astrocytoma"] = "Yes"
cancer2[cancer2 == "normal"] = "No"
cancer = factor(cancer, levels = c("Yes", "No"))

brain_df <- cbind(cancer2, brain)

```

#### Data train/test data splits based on binary output *cancer2* for logistic classification:

```{r logisticclassdata_LR}
set.seed(100)

### stratified random sampling to split up the data 
### while keeping the class proportions for binary class cancer:

brain_train1_idx <- createDataPartition(brain_df$cancer2, p=0.8, list = FALSE)
brain_trainb <- brain_df[brain_train1_idx,]
brain_testb <- brain_df[-brain_train1_idx,]

```


#### BoxCox, Scale and Center Training data for Logistic Regression:

```{r scalecenter_multiclass_LR}


brain_processb <- preProcess(brain_trainb[,-c(1:3)],
                    method = c("BoxCox", "scale", "center"))
brain_trainLR <- predict(brain_processb, brain_trainb)

```

#### PCA to reduce correlation among predictors since Logistic Regression is highly sensitive to correlated and large number of predictors:

```{r pca_LR}

brain_pcaLR <- preProcess(brain_trainLR[,-c(1:3)],
                    method = c("pca"))

brain_trainLR_pca <- predict(brain_pcaLR, brain_trainLR)

```

#### Model Logistic Regression Model:

```{r LR_model}

ctrl_LRC_brain <- trainControl(method = "cv",
                     summaryFunction = twoClassSummary,
                     classProbs = TRUE,
                     savePredictions = TRUE)

set.seed(476)

LR_model_brain <- train(x = brain_trainLR_pca[,-c(1:3)], 
               y = brain_trainLR_pca$cancer2,
               method = "glm",
               metric = "ROC",
               trControl = ctrl_LRC_brain)
LR_model_brain
LR_model_brain$finalModel
lrCM_brain <- confusionMatrix(LR_model_brain, norm = "none")
lrCM_brain

### Logistic Regression Results on training dta:
# Sensitivity: .8
# ROC: .8844
# Cross Validated Accuracy: .8570
# warnings regarding convergence with 0 or 1 probabilities.

```


***

### Nearest Shrunken data prep, train, and test:

#### Data train/test data splits based on multi-class output *type* on dataframe **brain** for Nearest Shrunken classification:

```{r multiclassdatasplit}

### Stratified random sampling for multiclass output classes, type:
set.seed(100)
brain$type=as.factor(brain$type)

brain_train_idx <- createDataPartition(brain$type, p=0.8, list = FALSE)
brain_train1 <- brain[brain_train_idx,]
brain_test1 <- brain[-brain_train_idx,]


```

#### BoxCox, Scale and Center Training data:

```{r scalecenter_multiclass}


brain_process <- preProcess(brain_train1[,-c(1:2)],
                    method = c("BoxCox", "scale", "center"))
brain2_train <- predict(brain_process, brain_train1)

```


#### Train Nearest Shrunken Centroid Model:

```{r shrunken_centroids}
ctrl_brain2 <- trainControl(method = "cv",
                     classProbs = TRUE,
                     savePredictions = TRUE)

set.seed(476)
nsc_model_brain2 <- train(x = brain2_train[,-c(1:2)], 
                y = brain2_train$type,
                method = "pam",
                tuneGrid = data.frame(threshold = seq(0, 25, length = 30)),
                trControl = ctrl_brain2)
nsc_model_brain2

### RESULTS Nearest Shrunken Centriods Training:
#Accuracy: .9336
#Kappa: .91004
#Cross-validated Accuracy = 
#threshold = 0
#number predictors: 54675

```
#### Most important variables for the NSC model on brain training data:

```{r model_results_nsc}

ImpVar_NCS_brain2 <- varImp(nsc_model_brain2, scale = FALSE)
plot(ImpVar_NCS_brain2, top=5)

```

#### BoxCox, center and scale test data:

```{r prepare_test_data}

brain_process_test <- preProcess(brain_test1[,-c(1:2)],
                    method = c("BoxCox", "scale", "center"))
brain2_test <- predict(brain_process_test, brain_test1)


```

#### Run the NSC model on test data:

```{r shurnken_centoids_results}
brain2_nsc_pred <- predict(nsc_model_brain2, brain2_test[,-c(1:2)])

confusionMatrix(brain2_nsc_pred, brain2_test$type)

```

***

### Extra code like PCA/SVM modeling unverified:

```{r pca}

brain_pca <- preProcess(brain_train[,-c(1,2)],
                    method = c("pca"))

brain_train_pca <- predict(brain_pca, brain_train)

```

```{r crossvalidation}

# Create 10x cross validation folds:
brain_cv_idx <- createFolds(brain_train_pca$type, returnTrain = TRUE)
brain_ctrl <- trainControl(method = "cv", index = brain_cv_idx)

```



```{r moredataexplore3}

head(brain_train_pca)
dim(brain_train_pca)

```

```{r moredataexplore4}

hist(brain_train_pca$PC1)
hist(brain_train_pca$PC1)
```

```{r moredataexplore5}
brain_trainX <- brain_train_pca[,-c(1,2)]
brain_testX1 <- brain_test1[,-c(1,2)]

brain_trainY <- brain_train_pca[,2]
brain_testY1 <- brain_test1[,2]



brain_trainY = factor(brain_trainY)

	
brain_testY1 = factor(brain_testY1)

set.seed(100)
svmR_brain_model <- train(x = brain_trainX, y = brain_trainY,
                  method = "svmLinear",
                  tuneLength = 14,
                  trControl = brain_ctrl)
svmR_model
#plot(svmR_brain_model, scales = list(x = list(log = 2))) 
svmRImpB <- varImp(svmR_brain_model, scale = FALSE)
plot(svmRImpB, top = 25)
#svmR_model$finalModel
# RESULTS:
# optimal C=4
# epsilon=0.1
# sigma=0.01826
# 1.329 RMSE, .4649 R^2

```



```{r plotsdata}



```


